Multi-Agent Path Finding with Matching using Increasing Cost Tree Search

### 2021-06-17 docs/paper.tex
Fixed grammar/spelling issues

	As was shown by Geiger et al.\cite{geiger2018}, in practical scenarios, only heuristic methods can solve TUSS instances. 
	Such methods are suboptimal and heuristic solutions to instances defined for a set shunting yard can be said to characterise a lower bound on the capacity of the shunting yard, as each solution cost gives a (loose) upper bound for the optimal solution cost. 
	The railway company, however, is interested in learning tight upper bounds for the capacity of existing shunting yards to inform decisions about matters like infrastructure expansion. 
	Mulderij et al. \cite{mulderij2020} propose the multi-agent pathfinding (MAPF) problem \cite{stern2019} extended with a matching subproblem (hereafter MAPFM) as a suitable relaxation for finding such upper-bounds using an approach analogous to that outlined above.
	For this to succeed, a way to optimally solve MAPFM instances is needed.

@@

	\begin{wrapfigure}[19]{R}{.25\textwidth}
		\begin{minipage}{\linewidth}
			\centering\captionsetup[subfigure]{justification=centering}
			\includegraphics[width=\linewidth]{img/vertex-conflict}
			\subcaption{Vertex conflict}
			\label{fig:conflictsa}\par\vfill
			\includegraphics[width=0.75\linewidth]{img/edge-conflict}
			\subcaption{Edge conflict}
			\label{fig:conflictsb}
		\end{minipage}
		\vspace{-10pt}
		\caption{The two types of conflict (from Fig. 1 in \cite{stern2019})}\label{fig:conflicts}
	\end{wrapfigure}
	This work is the outcome of a search for algorithms for optimally solving MAPFM based on ICTS as an alternative to the approach outlined above\footnote{To the knowledge of the author, to date, no such algorithm is described in the literature}, which is part of a project in which  other algorithms addition to ICTS were taken as starting points for MAPFM algorithms. Two classes of ICTS-based algorithms were found: one based on exhaustive enumeration of matchings using a variant of ICTS as subroutine, the other based on modifications to the ICT structure and its search. One exhaustive ICTS algorithm is compared to algorithms based on CBM\cite{ma2016}, M*\cite{wagner2011}, A*+OD+ID\cite{standley2010} and EPEA*\cite{goldenberg2014} respectively to put ICTS methods for MAPFM into a broader context.

@@

	As discussed in \cite{ma2016}, MAPFM generalizes both MAPF and anonymous MAPF, which is MAPF allowing an agent to move to any goal, a problem which can be solved in polynomial time.
	\section{Solving MAPF with Matching using ICTS} % Your contribution
	\label{section:icts-matching}
	Taking ICTS for MAPF as starting point, two methods to use ICTS for MAPFM were identified. The first method relies on a reduction from MAPFM to repeated MAPF: an optimal solution to MAPFM corresponds to a matching of agents to goals, so by exhaustively enumerating all matchings and solving these as MAPF instances, an optimal solution to the MAPFM instance can be found. In the second method, the increasing cost trees search itself is modified to allow an agent $a_j^i$ to be matched to any $g_k^i$. In this section, first ICTS for MAPF and some common variants will be described. Next, derived MAPFM algorithms are introduced and properties of these algorithms and some variants are discussed. In Section \ref{experiments}, these will be compared experimentally.
	\subsection{Increasing Cost Tree Search}
	\label{icts}
	In Increasing Cost Tree Search\cite{sharon2011}, the two-step mapping from path combination to SIC described in Section \ref{mapfm} is reversed to search for a solution with minimal SIC. This is done by searching for cost vectors corresponding to increasing SIC on the top level and searching for non-conflicting path combinations (solutions) for each cost vector on the bottom level.

@@

	The full ICT-search algorithm including a pruning step is described by Algorithm \ref{icts-algo}. It is reproduced from \cite{sharon2011} for completeness. As noted before, this procedure can be used for planning groups in Algorithm \ref{id-algo}. In this work, triple-pruning was used and enhanced triple-pruning was taken as a base-line, as the conclusion in \cite{sharon2011} is that this often outperforms other variants.
	\begin{algorithm}
		\begin{algorithmic}[1]
			\Procedure{ICT-search}{MAPF instance}
			\State \text{Construct ICT root}
			\ForAll{\text{ICT nodes }$(c_1,\ldots,c_k)$\text{ in breadth-first order }} 
			\ForAll{\text{agents} $a_i$} 
			\text{Construct} $MDD_i^{c_i}$
			\EndFor
			\ForAll{$j$-\text{agent combinations} $(i_1,\ldots,i_j)$} 
			\State \text{Run node-pruning procedure with MDDs }$MDD_{i_1}^{c_{i_1}},\ldots,MDD_{i_j}^{c_{i_j}}$
			\If{\text{no solution was found for a combination in pruning}}
			\State \textbf{Break}\text{ and continue with next ICT node}
			\EndIf
			\EndFor
			\State \text{Search }$k$\text{-agent MDD space}
			\If{\text{solution was found}}
			\State \textbf{return solution}
			\EndIf
			\EndFor

			\EndProcedure
		\end{algorithmic}
		\caption{Increasing Cost Tree Search}
		\label{icts-algo}
	\end{algorithm}
	\subsection{Exhaustive ICTS}
	\label{exhaustive}
	For $K$ teams $t_1,\ldots,t_K$ of size $k_i$, there are in total $M = \prod_{i} k_i!$ matchings from team agents to team goals. Let $C_1$ through $C_M$ denote the optimal SIC costs of each MAPF instance corresponding to the instance, with $C_m = \infty$ for unfeasible $m$. In order to find the optimal solution, $m$ with lowest $C_m$ has to be found, which requires all $M$ to be solved or found to be unfeasible, meaning that $\mathcal{O}(M)$ MAPF instances have to be solved. 
	\subsubsection{Bounded search}
	Once the first solution is found with cost $C$, this can be used in the next search as an (exclusive) upper-bound $B$ for the SIC of nodes that have to be searched. Formally, let $B_0 = \infty$ represent the initial SIC bound of the ICT search. After solving each $m$ with solution cost $C_m$, let $B_{m+1} = \min(B_m,C_m)$. By successively updating this $B$ in this manner, as more matchings are solved, the effort needed to solve the remaining matchings can be expected to be reduced significantly. With $C_m$ denoting the optimal SIC for matching $m$, this effort can be expressed as	
	\[\Delta_{m,bounded} = \min(B_m,C_m) - C^*_m = B_{m+1} - C^*_m\]
	This is proportional to the effort due to the complexity of the ICTS search for matching $m$, $\mathcal{O}(k^{\Delta_{m,bounded}})$.
	If $B_m < C^*_m$, only the root node will be generated to determine this so $\Delta_{m,bounded} = 0$.
	Using this, the number of ICT nodes searched in total becomes $\mathcal{O}(\sum_m k^{\Delta_{m,bounded}})$.
	\subsubsection{Ordered enumeration}
	\label{ordered-enum}
	In an effort to reduce the total effort as much as possible, the matchings $m$ can be enumerated and solved in increasing order of $C^*_m$. In this case, the effort per matching is monotonically non-increasing by Property \ref{monotonicity}. This induces a case in which this enumeration minimizes the number of nodes searched in total (Property \ref{cond-monoton-optim}). However, cases in which $\Delta_m$ of the first $m$'s searched is much higher could also be constructed, in which case this approach might cause more nodes to be processed in total than alternative methods. Therefore, this ordered enumeration remains a heuristic improvement that works well in practice, as shown in Section \ref{experiments}.

@@

	In implementation, ordered enumeration is approximated using a priority queue that maintains a subset of matchings of a fixed size $M_{pq}$\footnote{$M_{pq} = 10000$ was used in implementation but this could be much increased at the cost of memory}. Algorithm \ref{pq-algo} outlines the procedure.

@@

	%To avoid this, the MDD search could be changed to start at the maximal depth and search for a way towards the start node, which by start uniqueness remains without vertex conflicts.
	%Two strategies could be used to avoid this: firstly, the MDD search could be changed to start at the maximal depth and search for a way towards the start (which by start uniqueness is always non-conflicting); secondly, a bipartite matching algorithm could be employed to see if any matching exists between agents and goals they can reach. An advantage of the former strategy is that it will, in particular with larger team sizes, reduce the number of maximal depth nodes drastically, so that the subsequent search is from few nodes to one goal. By only running a matching algorithm to skip regular MDD search, there is no such reduction so from one starting position paths will be searched to potentially an exponential number of goal nodes, of which many will end up being illegal. 
	%[TODO: just focus on explaining why inverted MDD search was/could be used to solve this? Or just toss this out for lack of time/space. Why didn't I come up with this idea earlier...]
	\subsection{Optimizations}
	To improve both types of algorithms, some optimizations were found that can be used to speed up the search. These are here motivated. 
	\subsubsection{Bounds in ID}
	Information gained about solution costs of independent groups can be used to establish a lower bound on solution SIC as per Property \ref{property:id-lower}. In bounded search the upper bound can be further constrained by Property \ref{property:id-upper}.
	\begin{property}[Lower bound]
		When merging groups $g_i,g_j$ with solution costs $C_i,C_j$ to yield group $g_k$, for any solution for this group with cost $C_k$
		\[C_k\geq C_i + C_j\]
		\label{property:id-lower}
	\end{property}
	\begin{proof}
		Assume there would be a solution with $C_k < C_i + C_j$. Taking $C_k^i$ to denote the SIC of agents that used to make up $g_i$ in the solution to $g_k$, without loss of generality $C_k^i < C_i$, which contradicts the optimality of the solution with cost $C_i$ for $g_i$ independent of $g_j$.
	\end{proof}
	\begin{property}[Reduced upper bound]
		In bounded exhaustive search, the sum of costs of groups not being merged can be subtracted from bound $B$ without losing optimality.
		\label{property:id-upper}
	\end{property}
	\subsubsection{Child pruning}
	In the original ICTS formulation all $k$-children are generated in every case because pruning only controls skipping the low-level search, assumed to be the most expensive operation. However, in simple pruning in any case, when there is a conflict between two or more agents at a given ICT node $(c_1,\ldots,c_k)$, in any solution, one or more of the path costs corresponding to these conflicting agents will have to be incremented. This can be explioted by pruning the node's children that do not correspond to the conflicting agents, of which there are often just 2 or 3, drastically improving the (average) branching factor as $k$ increases. By Property \ref{property:ec-incompat}, child pruning is not compatible with enhanced pruning.
	\begin{property}[Incompatibility with enhanced pruning]
		Enhanced pruning combined with child pruning is suboptimal due to the \textit{cascading effect} described in \cite{sharon2011}.
		\label{property:ec-incompat}
	\end{property}

@@

	All algorithms are implemented in Python 3 to avoid performance differences as often seen between interpreted versus compiled languages; however, CBM\cite{baauw2021} in experiment 2 does use a SSP solver implemented in C++ as a subroutine. Nonetheless, the comparison is here a bit more coarse-grained as performance differences are sometimes subtle and could potentially be explained by slight implementation optimizations rather than algorithmic properties.

@@

	Between exhaustive variants, with exception of the ID variant, a type of hierarchy can be distinguished, best illustrated by \ref{fig:i-25-3-p}; almost without exception, Exh+S+C+B+O outperforms Exh+E+B+O which outperforms Exh+E+B which in turn outperforms Exh+E. Bounded search outperforms unbounded search as expected, given that the former does at most as much work as the latter and often less. Approximate ordering outperforming unordered enumeration is not quite as obvious from a theoretical standpoint, as discussed in \ref{ordered-enum}, and simply indicates that sorting works well in practice. Simple pruning with child pruning outperforming exhaustive pruning shows the potential of this pruning method. In particular as $k$ grows larger and in scenarios with more obstacles (\ref{fig:i-75-1-p},\ref{fig:i-75-3-p}), it appears that having to search a smaller ICT is more important in terms of performance than being able to skip the $k$-agent low-level search more often. As discussed in Section \ref{future}, more work is needed to determine if such an improvement is to be expected more generally.

@@

	\subsection{Experiment 2: Algorithms for MAPFM}
	From the variants discussed and compared above, the Exh+E+B+O variant was selected for comparison with various other algorithms for MAPFM. This results in the following set of algorithms whose performance is shown in Figure \ref{fig:r-probs}.
	\begin{itemize}
		\begin{minipage}{0.45\linewidth}
			\item ICTS with approximate sorted exhaustive matching (Exh+E+B+O)
			\item A*+ID+OD with matching ID and sorted exhaustive matching\cite{bruin2021}
			\item EPEA* with matching ID and sorted exhaustive matching\cite{jong2021}
		\end{minipage}
		\hspace{1cm}
		\begin{minipage}{0.40\linewidth}
			\item M* with sorted exhaustive matching\cite{donszelmann2021}
			\item CBM adapted to use SIC, using the successive shortest path algorithm (SSP) \cite{goldberg1987} as subroutine for solving teams\cite{baauw2021}.
		\end{minipage}
	\end{itemize}
	With exception of CBM, all methods described use a variant of sorted exhaustive matching. This is due to the fact that from each family of algorithms, one of the best performing algorithms was chosen, which always was such an exhaustive method as opposed to a generalization of the base algorithm to MAPFM (ICTS-m in this work). A*+ID+OD and EPEA* also make use of matching ID, which is ID applied to the MAPFM instance using the respective exhaustive algorithm as MAPFM solver. Hypothetically, this could explain why both outperform the methods using just sorted exhaustive matching in maps with 75\% wall and 3 teams (\ref{fig:r-75-3-p}); this difference is not quite as clear in the remaining scenarios shown in Figure \ref{fig:r-probs} however.

	CBM solving all problems with one team (Figure \ref{fig:r-25-1-p} and \ref{fig:r-75-1-p}), problems equivalent to anonymous MAPF instances, is at first surprising, but follows directly from the definition of the CBM method. CBM solves single teams as anonymous MAPF instances in polynomial time, meaning that solving a single team reduces to a polynomial-time procedure, namely SSP. It is expected that conform the results reported in \cite{ma2016}, as the number of teams approaches $k$, CBM performance will deteriorate and thus become more similar to the results for exhaustive methods here reported.
%	Still, some implementations might be more optimized than others so this might be a factor in performance characteristics and in particular in relative performance differences.
	\section{Conclusions and future work}
	\label{conclusions}
	\label{future}
	The goal of this research was to seek alternative approaches to solving MAPFM based on ICTS. In this, the research has succeeded, but the scalability of the algorithms leaves something to be desired. Exhaustive methods, like those discussed in detail in \ref{exhaustive}, cannot be expected to scale well as teams grow larger due to their complexity, no matter what optimizations are put in place. Nonetheless, as was shown in Section \ref{experiments}, the more advanced exhaustive ICTS algorithms consistently outperform the ICTS-m based algorithms. Compared to the SIC-adaptation of the CBM method \cite{baauw2021}, it was seen that exhaustive methods in general perform quite poorly if there are fewer teams and fewer obstacles, while being competitive once teams grow smaller and maps become more constrained. Delineating under which conditions exhaustive methods outperform CBM and vice versa is something for further research.

@@

	\section{Responsible Research}
	The implementation of the ICTS-based algorithms described in this work is publicly available online in the form of a Github repository\footnote{\url{https://github.com/tbvanderwoude/icts-m}}. Included is also code for benchmarking different configurations of the ICTS solver and the maps that were used in the benchmarks reported above. The benchmarks and maps used in relative comparison of MAPFM algorithms are also available in the code and data repository of \cite{donszelmann2021}\footnote{\url{https://github.com/jonay2000/research-project}}.

	\printbibliography
	\newgeometry{left=1cm,right=1cm,top=0.1cm,bottom=0.1cm}
	\begin{figure*}[t]
		\centering
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/25-1}
			\caption{25\% wall, 1 team}
			\label{fig:i-25-1}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/25-3}
			\caption{25\% wall, 3 teams}
			\label{fig:i-25-3}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/75-1}
			\caption{75\% wall, 1 team}
			\label{fig:i-75-1}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/75-3}
			\caption{75\% wall, 3 teams}
			\label{fig:i-75-3}
		\end{subfigure}
		\caption{Average solving time of different ICTS-based algorithms with a timeout of 120s}
		\label{fig:i-times}
	\end{figure*}
	\begin{figure*}[b]
		\centering
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/25-1}
			\caption{25\% wall, 1 team}
			\label{fig:r-25-1}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/25-3}
			\caption{25\% wall, 3 teams}
			\label{fig:r-25-3}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/75-1}
			\caption{75\% wall, 1 team}
			\label{fig:r-75-1}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/75-3}
			\caption{75\% wall, 3 teams}
			\label{fig:r-75-3}
		\end{subfigure}
		\caption{Average solving time of MAPFM algorithms timeout of 120s}
		\label{fig:r-times}
	\end{figure*}

>>>>

	As was shown by Geiger et al.\cite{geiger2018}, in practical scenarios, only heuristic methods can solve TUSS instances. 
	Such methods are suboptimal and heuristic solutions to instances defined for a set shunting yard can be said to characterise a lower bound on the capacity of the shunting yard, as each solution cost gives a (loose) upper bound for the optimal solution cost. 
	The railway company, however, is interested in learning tight upper bounds for the capacity of existing shunting yards to inform decisions about matters like infrastructure expansion. 
	Mulderij et al. \cite{mulderij2020} propose the multi-agent pathfinding (MAPF) problem \cite{stern2019} extended with a matching subproblem (hereafter MAPFM) as a suitable relaxation of (discretised) TUSS for finding such upper-bounds using an approach analogous to that outlined above.
	For this to succeed, a way to optimally solve MAPFM instances is needed.

@@

	\begin{wrapfigure}[19]{R}{.25\textwidth}
		\begin{minipage}{\linewidth}
			\centering\captionsetup[subfigure]{justification=centering}
			\includegraphics[width=\linewidth]{img/vertex-conflict}
			\subcaption{Vertex conflict}
			\label{fig:conflictsa}\par\vfill
			\includegraphics[width=0.75\linewidth]{img/edge-conflict}
			\subcaption{Edge conflict}
			\label{fig:conflictsb}
		\end{minipage}
		\vspace{-10pt}
		\caption{The two types of conflict (from Fig. 1 in \cite{stern2019})}\label{fig:conflicts}
	\end{wrapfigure}
	This work is the outcome of a search for algorithms for optimally solving MAPFM based on ICTS as an alternative to the approach outlined above\footnote{To the knowledge of the author, to date, no such algorithm is described in the literature}, which is part of a project in which other algorithms addition to ICTS were taken as starting points for MAPFM algorithms. Two classes of ICTS-based algorithms were found: one based on exhaustive enumeration of matchings using a variant of ICTS as a subroutine, the other based on modifications to the ICT structure and its search. One exhaustive ICTS algorithm is compared to algorithms based on CBM\cite{ma2016}, M*\cite{wagner2011}, A*+OD+ID\cite{standley2010} and EPEA*\cite{goldenberg2014} respectively to put ICTS methods for MAPFM into a broader context.

@@

	As discussed in \cite{ma2016}, MAPFM generalizes both MAPF and anonymous MAPF, which is MAPF allowing an agent to move to any goal, a problem that can be solved in polynomial time.
	\section{Solving MAPF with Matching using ICTS} % Your contribution
	\label{section:icts-matching}
	Taking ICTS for MAPF as starting point, two methods to use ICTS for MAPFM were identified. The first method relies on a reduction from MAPFM to repeated MAPF: an optimal solution to MAPFM corresponds to a matching of agents to goals, so by exhaustively enumerating all matchings and solving these as MAPF instances, an optimal solution to the MAPFM instance can be found. In the second method, the increasing cost trees search itself is modified to allow an agent $a_j^i$ to be matched to any $g_k^i$. First, ICTS for MAPF and some common variants will be described. Next, derived MAPFM algorithms are introduced and properties of these algorithms and some variants are discussed. In Section \ref{experiments}, these will be compared experimentally.
	\subsection{Increasing Cost Tree Search}
	\label{icts}
	In Increasing Cost Tree Search\cite{sharon2011}, the two-step mapping from path combination to SIC described in Section \ref{mapfm} is reversed to search for a solution with minimal SIC. This is done by searching for cost vectors corresponding to increasing SIC on the top level and searching for non-conflicting path combinations (solutions) for each cost vector on the bottom level.

@@

	The full ICT-search algorithm including a pruning step is described by Algorithm \ref{icts-algo}. It is reproduced from \cite{sharon2011} for completeness. As noted before, this procedure can be used for planning groups in Algorithm \ref{id-algo}. In this work, triple-pruning was used and enhanced triple-pruning was taken as a baseline, as the conclusion in \cite{sharon2011} is that this often outperforms other variants.
	\begin{algorithm}
		\begin{algorithmic}[1]
			\Procedure{ICT-search}{MAPF instance}
			\State \text{Construct ICT root}
			\ForAll{\text{ICT nodes }$(c_1,\ldots,c_k)$\text{ in breadth-first order }} 
			\ForAll{\text{agents} $a_i$} 
			\text{Construct} $MDD_i^{c_i}$
			\EndFor
			\ForAll{$j$-\text{agent combinations} $(i_1,\ldots,i_j)$} 
			\State \text{Run node-pruning procedure with MDDs }$MDD_{i_1}^{c_{i_1}},\ldots,MDD_{i_j}^{c_{i_j}}$
			\If{\text{no solution was found for a combination in pruning}}
			\State \textbf{Break}\text{ and continue with next ICT node}
			\EndIf
			\EndFor
			\State \text{Search }$k$\text{-agent MDD space}
			\If{\text{solution was found}}
			\State \textbf{return solution}
			\EndIf
			\EndFor

			\EndProcedure
		\end{algorithmic}
		\caption{Increasing Cost Tree Search}
		\label{icts-algo}
	\end{algorithm}
	\subsection{Exhaustive ICTS}
	\label{exhaustive}
	For $K$ teams $t_1,\ldots,t_K$ of size $k_i$, there are in total $M = \prod_{i} k_i!$ matchings from team agents to team goals. Let $C_1$ through $C_M$ denote the optimal SIC costs of each MAPF instance corresponding to the instance, with $C_m = \infty$ for unfeasible $m$. In order to find the optimal solution, $m$ with lowest $C_m$ has to be found, which requires all $M$ to be solved or found to be unfeasible, meaning that $\mathcal{O}(M)$ MAPF instances have to be solved. 
	\subsubsection{Bounded search}
	Once the first solution is found with cost $C$, this can be used in the next search as an (exclusive) upper-bound $B$ for the SIC of nodes that have to be searched. Formally, let $B_0 = \infty$ represent the initial SIC bound of the ICT search. After solving each $m$ with solution cost $C_m$, let $B_{m+1} = \min(B_m,C_m)$. By successively updating this $B$ in this manner, as more matchings are solved, the effort needed to solve the remaining matchings can be expected to be reduced significantly. With $C_m$ denoting the optimal SIC for matching $m$, this effort can be expressed as	
	\[\Delta_{m,bounded} = \min(B_m,C_m) - C^*_m = B_{m+1} - C^*_m\]
	This is proportional to the effort due to the complexity of the ICTS search for matching $m$, $\mathcal{O}(k^{\Delta_{m,bounded}})$.
	If $B_m < C^*_m$, only the root node will be generated to determine this so $\Delta_{m,bounded} = 0$.
	Using this, the number of ICT nodes searched in total becomes $\mathcal{O}(\sum_m k^{\Delta_{m,bounded}})$.
	\subsubsection{Ordered enumeration}
	\label{ordered-enum}
	To reduce the total effort as much as possible, the matchings $m$ can be enumerated and solved in increasing order of $C^*_m$. In this case, the effort per matching is monotonically non-increasing by Property \ref{monotonicity}. This induces a case in which this enumeration minimizes the number of nodes searched in total (Property \ref{cond-monoton-optim}). However, cases in which $\Delta_m$ of the first $m$'s searched is much higher could also be constructed, in which case this approach might cause more nodes to be processed in total than alternative methods. Therefore, this ordered enumeration remains a heuristic improvement that works well in practice, as shown in Section \ref{experiments}.

@@

	In the implementation, ordered enumeration is approximated using a priority queue that maintains a subset of matchings of a fixed size $M_{pq}$\footnote{$M_{pq} = 10000$ was used in implementation but this could be much increased at the cost of memory}. Algorithm \ref{pq-algo} outlines the procedure.

@@

	%To avoid this, the MDD search could be changed to start at the maximal depth and search for a way towards the start node, which by start uniqueness remains without vertex conflicts.
	%Two strategies could be used to avoid this: firstly, the MDD search could be changed to start at the maximal depth and search for a way towards the start (which by start uniqueness is always non-conflicting); secondly, a bipartite matching algorithm could be employed to see if any matching exists between agents and goals they can reach. An advantage of the former strategy is that it will, in particular with larger team sizes, reduce the number of maximal depth nodes drastically, so that the subsequent search is from few nodes to one goal. By only running a matching algorithm to skip regular MDD search, there is no such reduction so from one starting position paths will be searched to potentially an exponential number of goal nodes, of which many will end up being illegal. 
	%[TODO: just focus on explaining why inverted MDD search was/could be used to solve this? Or just toss this out for lack of time/space. Why didn't I come up with this idea earlier...]
	\subsection{Optimizations}
	To improve both types of algorithms, some optimizations were found that can be used to speed up the search. These are here motivated. 
	\subsubsection{Bounds in ID}
	The information gained about solution costs of independent groups can be used to establish a lower bound on solution SIC as per Property \ref{property:id-lower}. In bounded search, the upper bound can be further constrained by Property \ref{property:id-upper}.
	\begin{property}[Lower bound]
		When merging groups $g_i,g_j$ with solution costs $C_i,C_j$ to yield group $g_k$, for any solution for this group with cost $C_k$
		\[C_k\geq C_i + C_j\]
		\label{property:id-lower}
	\end{property}
	\begin{proof}
		Assume there would be a solution with $C_k < C_i + C_j$. Taking $C_k^i$ to denote the SIC of agents that used to make up $g_i$ in the solution to $g_k$, without loss of generality $C_k^i < C_i$, which contradicts the optimality of the solution with cost $C_i$ for $g_i$ independent of $g_j$.
	\end{proof}
	\begin{property}[Reduced upper bound]
		In bounded exhaustive search, the sum of costs of groups not being merged can be subtracted from bound $B$ without losing optimality.
		\label{property:id-upper}
	\end{property}
	\subsubsection{Child pruning}
	In the original ICTS formulation all $k$-children are generated in every case because pruning only controls skipping the low-level search, assumed to be the most expensive operation. However, in simple pruning in any case, when there is a conflict between two or more agents at a given ICT node $(c_1,\ldots,c_k)$, in any solution, one or more of the path costs corresponding to these conflicting agents will have to be incremented. This can be exploited by pruning the node's children that do not correspond to the conflicting agents, of which there are often just 2 or 3, drastically improving the (average) branching factor as $k$ increases. By Property \ref{property:ec-incompat}, child pruning is not compatible with enhanced pruning.
	\begin{property}[Incompatibility with enhanced pruning]
		Enhanced pruning combined with child pruning is suboptimal due to the \textit{cascading effect} described in \cite{sharon2011}.
		\label{property:ec-incompat}
	\end{property}

@@

	All algorithms are implemented in Python 3 to avoid performance differences as often seen between interpreted versus compiled languages; however, CBM\cite{baauw2021} in experiment 2 does use an SSP solver implemented in C++ as a subroutine. Nonetheless, the comparison is here a bit more coarse-grained as performance differences are sometimes subtle and could potentially be explained by slight implementation optimizations rather than algorithmic properties.

@@

	Between exhaustive variants, with exception of the ID variant, a type of hierarchy can be distinguished, best illustrated by \ref{fig:i-25-3-p}; almost without exception, Exh+S+C+B+O outperforms Exh+E+B+O which outperforms Exh+E+B which in turn outperforms Exh+E. Bounded search outperforms unbounded search as expected, given that the former does at most as much work as the latter and often less. Approximate ordering outperforming unordered enumeration is not quite as obvious from a theoretical standpoint, as discussed in \ref{ordered-enum}, and simply indicates that sorting works well in practice. Simple pruning with child pruning outperforming exhaustive pruning shows the potential of this pruning method. In particular, as $k$ grows larger and in scenarios with more obstacles (\ref{fig:i-75-1-p},\ref{fig:i-75-3-p}), it appears that having to search a smaller ICT is more important in terms of performance than being able to skip the $k$-agent low-level search more often. As discussed in Section \ref{future}, more work is needed to determine if such an improvement is to be expected more generally.

@@

	\subsection{Experiment 2: Algorithms for MAPFM}
	From the variants discussed and compared above, the Exh+E+B+O variant was selected for comparison with various other algorithms for MAPFM. This results in the following set of algorithms whose performance is shown in Figure \ref{fig:r-probs}.
	\begin{itemize}
		\begin{minipage}{0.45\linewidth}
			\item ICTS with approximate sorted exhaustive matching (Exh+E+B+O)
			\item A*+ID+OD with matching ID and sorted exhaustive matching\cite{bruin2021}
			\item EPEA* with matching ID and sorted exhaustive matching\cite{jong2021}
		\end{minipage}
		\hspace{1cm}
		\begin{minipage}{0.40\linewidth}
			\item M* with sorted exhaustive matching\cite{donszelmann2021}
			\item CBM adapted to use SIC, using the successive shortest path algorithm (SSP) \cite{goldberg1987} as subroutine for solving teams\cite{baauw2021}.
		\end{minipage}
	\end{itemize}
	With exception of CBM, all methods described use a variant of sorted exhaustive matching. This is because from each family of algorithms one of the best performing algorithms was chosen, which always was such an exhaustive method as opposed to a generalization of the base algorithm to MAPFM (ICTS-m in this work). A*+ID+OD and EPEA* also make use of matching ID, which is ID applied to the MAPFM instance using the respective exhaustive algorithm as MAPFM solver. Hypothetically, this could explain why both outperform the methods using just sorted exhaustive matching in maps with 75\% wall and 3 teams (\ref{fig:r-75-3-p}); this difference is not quite as clear in the remaining scenarios shown in Figure \ref{fig:r-probs} however.

	CBM solving all problems with one team (Figure \ref{fig:r-25-1-p} and \ref{fig:r-75-1-p}), problems equivalent to anonymous MAPF instances, is at first surprising but follows directly from the definition of the CBM method. CBM solves single teams as anonymous MAPF instances in polynomial time, meaning that solving a single team reduces to a polynomial-time procedure, namely SSP. Following the results reported in \cite{ma2016}, it is expected that as the number of teams approaches $k$, CBM performance will deteriorate and thus become more similar to the results for exhaustive methods here reported.
	%	Still, some implementations might be more optimized than others so this might be a factor in performance characteristics and in particular in relative performance differences.
	\section{Conclusions and Future work}
	\label{conclusions}
	\label{future}
	The goal of this research was to seek alternative approaches to solving MAPFM based on ICTS. In this, the research has succeeded, but the scalability of the algorithms leaves something to be desired. Exhaustive methods, like those discussed in detail in \ref{exhaustive}, cannot be expected to scale well as teams grow larger due to their complexity, no matter what optimizations are put in place. Nonetheless, as was shown in Section \ref{experiments}, the more advanced exhaustive ICTS algorithms consistently outperform the ICTS-m based algorithms. Compared to the SIC-adaptation of the CBM method \cite{baauw2021}, it was seen that exhaustive methods generally perform quite poorly if there are fewer teams and fewer obstacles, while being competitive once teams grow smaller and maps become more constrained. Delineating under which conditions exhaustive methods outperform CBM and vice versa is something for further research.

@@

	\section{Responsible Research}
	The implementation of the ICTS-based algorithms described in this work is publicly available online in the form of a Github repository\footnote{\url{https://github.com/tbvanderwoude/icts-m}}. Included is also code for benchmarking different configurations of the ICTS solver and the maps that were used in the benchmarks reported above. The benchmarks and maps used in the relative comparison of MAPFM algorithms are also available in the code and data repository of \cite{donszelmann2021}\footnote{\url{https://github.com/jonay2000/research-project}}.

	\printbibliography
	\newgeometry{left=1cm,right=1cm,top=0.1cm,bottom=0.1cm}
	\begin{figure*}[t]
		\centering
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/25-1}
			\caption{25\% wall, 1 team}
			\label{fig:i-25-1}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/25-3}
			\caption{25\% wall, 3 teams}
			\label{fig:i-25-3}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/75-1}
			\caption{75\% wall, 1 team}
			\label{fig:i-75-1}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/75-3}
			\caption{75\% wall, 3 teams}
			\label{fig:i-75-3}
		\end{subfigure}
		\caption{Average solving time of different ICTS-based algorithms with a timeout of 120s}
		\label{fig:i-times}
	\end{figure*}
	\begin{figure*}[b]
		\centering
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/25-1}
			\caption{25\% wall, 1 team}
			\label{fig:r-25-1}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/25-3}
			\caption{25\% wall, 3 teams}
			\label{fig:r-25-3}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/75-1}
			\caption{75\% wall, 1 team}
			\label{fig:r-75-1}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/75-3}
			\caption{75\% wall, 3 teams}
			\label{fig:r-75-3}
		\end{subfigure}
		\caption{Average solving time of MAPFM algorithms with a timeout of 120s}
		\label{fig:r-times}
	\end{figure*}
